Jayden Stats Homework

### 2025-06-18 DeficiencyScreen/Prelim_Analysis.Rmd
updated df data

genesindf = list()
i=1
while(i <= nrow(fullDfKey)){
  temp = na.omit(gene.location[((gl <= Dfl[i] & gr >= Dfr[i]) |
                               (gl >= Dfl[i] & gl <= Dfr[i]) |
                               (gr >= Dfl[i] & gr <= Dfr[i])) &
                               gene.location$CHR == fullDfKey$Chr[i],])

  genesindf[[fullDfKey$Deficiency[i]]] = temp$FBgn
  i=i+1
}

>>>>

genesindf = data.frame(Df = character(),
                       FBgn = character(),
                       Symbol = character())
i=1
while(i <= nrow(fullDfKey)){
  temp = na.omit(gene.location[((gl <= Dfl[i] & gr >= Dfr[i]) |
                               (gl >= Dfl[i] & gl <= Dfr[i]) |
                               (gr >= Dfl[i] & gr <= Dfr[i])) &
                               gene.location$CHR == fullDfKey$Chr[i],])

  if(nrow(temp) > 0){
    temp$Df = fullDfKey$Deficiency[i]
    genesindf = rbind(genesindf, temp[, c("Df", "FBgn", "Symbol")])
  }
  #genesindf[[fullDfKey$Deficiency[i]]] = temp$FBgn

  i=i+1
}

python.genesindf = read.csv("/Users/johncsantiago/Documents/GitHub/Python/DeficiencyData/Output.csv", row.names = 1)

py.gid = python.genesindf[,c(3,1,2)]
colnames(py.gid) = c("Df", "FBgn", "Symbol")
py.gid$Df = fullDfKey[py.gid$Df, "Deficiency"]

##Dealing with nomenclature issues
##identifying genes with an FBgn found using python method that were not identified using org.Dm.eg.db approach

py.only = py.gid[py.gid$FBgn %in% setdiff(py.gid$FBgn, gene.location$FBgn),]

##of these 23 genes, how many have a symbol only found with python

py.only = py.only[py.only$Symbol %in% setdiff(py.only$Symbol, gene.location$Symbol),]

##only 8 genes that were not found using the org.Dm.eg.db approach by symbol or FBgn

##unique(py.only$Symbol)
#"Est17" found under CG10175 in R data
#"Dld" found under CG7430 in R data

##NOT in the R data
#"lncRNA:CR46532" 
#"asRNA:PS4"     
#"lncRNA:CR46530" 
#"lncRNA:CR46535"
#"lncRNA:CR46529"
#"lncRNA:CR46534"

py.only.FBgn = c("lncRNA:CR46532", "FBgn0288728", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.sym = c("lncRNA:CR46532", "asRNA:PS4", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.ens = c("lncRNA:CR46532", "asRNA:PS4", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.CHR = c("lncRNA:CR46532", "3R", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.Start = c("lncRNA:CR46532", 16720127, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.End = c("lncRNA:CR46532", 16729959, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.Strand = c("lncRNA:CR46532", 1, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

add.gene.locations = data.frame(FBgn = py.only.FBgn,
                                Symbol = py.only.sym,
                                ensembl = py.only.ens,
                                CHR = py.only.CHR,
                                Start = py.only.Start,
                                End = py.only.End,
                                Strand = py.only.Strand)

### 2025-06-18 GOandKEGGTermEnrichment.Rmd
Updated Russo Data
updated df data

Variable.Condition = "GRFC" 
Control.Condition = "WTFC"
secondary.Variable.Condition = "GRMC"
secondary.Control.Condition = "WTMC"

@@

@@

##can be the exact name or 5 digit code in quotes
specific.kegg = "00980"

@@

##5 digits from KEGG ID
PathwayID = "00010"

@@

##can be the exact name or 5 digit code in quotes
specific.kegg = "00010"

@@

@@

##a parent category generated in the treemapPlot function
parent.category = 'membrane lipid biosynthetic process'

>>>>

#A4V conditions use 4 parameters:   A or S for genotype
#                                   3, 9 or 40 for time
#                                   M or F for sex
#                                   H, T or A for body section
#                                   example: A3MT

#G85R conditions use 3 parameters:  GR or WT for genotype
#                                   M or F for sex
#                                   C, Df or OE for TKT modulation
#                                   example: GRMOE
Variable.Condition = "GRFC" 
Control.Condition = "WTFC"
secondary.Variable.Condition = "GRMC"
secondary.Control.Condition = "WTMC"

@@

cat.data = Subset.Enrichment(row.names(FDR), row.names(FDR)[FDR[,1]<=0.05 & FDR[,2]<=0.05])

cat.data = Subset.Enrichment(row.names(FDR), row.names(FDR)[FDR[,1]>0.05 & FDR[,2]<=0.05])

cat.data = Subset.Enrichment(row.names(FDR), row.names(FDR)[FDR[,1]<=0.05 & FDR[,2]>0.05])

@@

##can be the exact name or 5 digit code in quotes
specific.kegg = "04146"

@@

##5 digits from KEGG ID
PathwayID = "03050"

@@

##can be the exact name or 5 digit code in quotes
specific.kegg = "03050"

@@

##the 7 digits of a GOterm id or the exact term name
specific.go = '0043228'

@@

##a parent category generated in the treemapPlot function
parent.category = 'double-strand break repair via break-induced replication'

### 2025-06-18 GeneralDataFiles/CompleteMetabolicNetwork.Rmd
updated df data

G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)

raw.nodes = read.csv(paste0(git.dir, "combined.nodes.csv"), row.names = 1)
raw.edges = read.csv(paste0(git.dir, "combined.edges.csv"))

temp = colorRampPalette(c('royalblue',"dodgerblue",'red', "firebrick"))(401)

@@

nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = 'dot'

node.colors = G85R.metabFC[nodes$data.id, "GR.CxOE"]
node.colors = temp[(200*node.colors/max(abs(na.exclude(node.colors)))) + 201]
node.colors[na.action(na.exclude(node.colors))] = "white"

nodes$color.background = node.colors

>>>>

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)
G85R.geneFDR = read.csv(paste0(git.dir, "TKT.EdgeR.FDRTable.csv"), row.names = 1)
G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)
G85R.metabFC[,2:ncol(G85R.metabFC)] = log2(G85R.metabFC[,2:ncol(G85R.metabFC)])
G85R.metabFDR = read.csv(paste0(git.dir, "MetaboliteFDRs.csv"), row.names = 1)

raw.nodes = read.csv(paste0(git.dir, "combined.nodes.csv"), row.names = 1)
raw.nodes$x = raw.nodes$x*1
raw.nodes$y = raw.nodes$y*1

raw.edges = read.csv(paste0(git.dir, "combined.edges.csv"))

temp = colorRampPalette(c("navy", "royalblue","white", "deeppink", "firebrick"))(401)
temp2 = colorRampPalette(c("turquoise", "grey", "deeppink"))(401)

##metabolomics Choices
##"GRxWT.C"  "GRxWT.Df" "GRxWT.OE" "GR.CxDf" "WT.CxDf"  "GR.CxOE"  "WT.CxOE"
metab = "GRxWT.C"

##Transcriptome choices
##"GRF.CxDf" "GRF.CxOE"  "WTF.CxDf" "WTF.CxOE"  "GRxWT.FC" "GRxWT.FDf"
##"GRxWT.FOE" "GRM.CxDf"  "GRM.CxOE" "WTM.CxDf"  "WTM.CxOE"  "GRxWT.MC"
##"GRxWT.MDf" "GRxWT.MOE" "GR.FxM" "WT.FxM" 
enzyme = "GRxWT.FDf"

nodes[nodes$shape == "dot", "label"] = NA

nodes.color = G85R.metabFC[nodes[nodes$shape == "circle", "data.id"], metab]
nodes.color = temp[((nodes.color/max(abs(na.omit(nodes.color))))*200)+201]
nodes[nodes$shape == "circle", "color.background"] = nodes.color
nodes$color.background[na.action(na.exclude(nodes$color.background))] = "white"

node.size = G85R.metabFDR[nodes[nodes$shape == "circle", "data.id"], metab]
node.size[na.action(na.exclude(node.size))] = 1
node.size = -log10(node.size)+1
node.size = node.size*10
node.size[node.size>80] = 50
node.size[node.size < (-log10(.2)+1)*10] =10
nodes[nodes$shape == "circle", "size"] = node.size

nodes$shape = 'dot'

nodes$title = paste0(nodes$id,
                     "<br>-log2(FC): ", 
                     signif((G85R.metabFC[nodes$data.id, metab]), 3),
                     "<br>FDR: ", 
                     signif((G85R.metabFDR[nodes$data.id, metab]), 3))

edges$title = paste0(GeneIDKey[edges$FBgn, "Symbol"],
                     "<br>-log2(FC): ", 
                     signif(G85R.geneFC[edges$FBgn,
                                 enzyme],3),
                     "<br>FDR: ", 
                     signif(G85R.geneFDR[edges$FBgn,
                                 enzyme],3))

edges.width = G85R.geneFDR[edges$FBgn, enzyme]
edges.width[na.action(na.exclude(edges.width))] = 1
edges.width = (-log10(edges.width))+1
edges.width = edges.width * 2

edges.width[edges.width>6] = 6
edges.width[edges.width < (-log10(.2)+1)*2] =1

edges$width = edges.width

edges.colors = G85R.geneFC[edges$FBgn, enzyme]
edges.colors = temp[((edges.colors/max(abs(na.omit(edges.colors))))*200)+201]
edges.colors[na.action(na.exclude(edges.colors))] = "grey"

edges$color = edges.colors

visNetwork(nodes, edges, background = "black")%>%
  visNodes(physics = F)%>%
  #visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = list(enabled = T, values = nodes[nodes$size > 0, "label"]))%>%
  visNodes(color = list(border = "white"), font = list(color = "gold", size = 14, background = "black"))%>%
  visEdges(arrows = visedges$arrows, physics = T)

@@

nodes[nodes$shape == "dot", "label"] = NA
nodes[nodes$shape == "circle", "size"] = 20
nodes.color = G85R.metabFC[nodes[nodes$shape == "circle", "data.id"], metab]
nodes.color = temp[((nodes.color/max(abs(na.omit(nodes.color))))*200)+201]
nodes[nodes$shape == "circle", "color.background"] = nodes.color
nodes$color.background[na.action(na.exclude(nodes$color.background))] = "ghostwhite"
nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = "dot"

edges$color = "black"

PPP.genes = c("FBgn0023477",
              "FBgn0050499",
              "FBgn0050410",
              "FBgn0004654",
              "FBgn0030239",
              "FBgn0037147",
              "FBgn0004057",
              "FBgn0037607",
              "FBgn0036784")

edges[edges$FBgn %in% c(PPP.genes), "color"] = "black"
edges[edges$FBgn %in% c(PPP.genes), "width"] = 2.5
#edges[ grep("FBgn0036784", edges$FBgn), "color"] = "darkgreen"
#edges[ grep("FBgn0037607", edges$FBgn), "color"] = "darkorange"

visNetwork(nodes, edges)%>%
  visNodes(physics = F, font = list(size = 18, background = hex("white", .6)))

#edges[grep('red', edges$color),]

### 2025-06-18 RussoPaper/Prelim_UBLegMorphologyDataAnalysis.Rmd
updated df data

heat.data = temp[,6:14]
row.names(heat.data) = paste0(temp$Genotype,temp$Animal..)
heat.data = as.matrix(t(heat.data))

@@

heatmap.2(heat.data,
          Colv = T,
          Rowv = T,
          margins = c(7, 15),
          trace = "none",
          ColSideColors = sample.colors,
          RowSideColors = category.colors,
          col = heat.colors)
```

@@

@@

>>>>

heat.data = temp[,6:14]
heat.data = temp[,c(7:9,12:14)]

row.names(heat.data) = paste0(temp$Genotype,temp$Animal..)
heat.data = as.matrix(t(heat.data))

@@

heatmap.2(heat.data,
          Colv = T,
          Rowv = F,
          dendrogram = "column",
          margins = c(7, 15),
          trace = "none",
          ColSideColors = sample.colors,
          #RowSideColors = category.colors,
          col = heat.colors)
```

@@

temp.combos$combo = (substr(temp.combos$combo, 1, nchar(temp.combos$combo)-2))
temp.combos= temp.combos[!duplicated(temp.combos[, "combo"]),]

temp.combos[1:20, 2:4]

@@

```{r}

library(ggridges)

#best.combo = data_matrix[,c(2,3,4,7,8,9)]

best.titles = temp.titles[c(2,3,4,7,8,9)]

morph.data = temp[,c(1:2,7:9,12:14)]

WT.morphdata = morph.data[morph.data$Genotype == "WT(cr)",]
WT.morphdata = t(WT.morphdata[,3:8])
WT.morphdata = as.data.frame(WT.morphdata)
WT.morphdata$Morph = best.titles
wt.morphplot = data.frame(Morph = as.character(),
                          Score = as.numeric())
i=1
while(i<=9){
  temp.morph = data.frame(Morph = WT.morphdata$Morph,
                          Score = WT.morphdata[,i])
  wt.morphplot = rbind(wt.morphplot, temp.morph)
  i = i +1
}
wt.morphplot$Morph = factor(wt.morphplot$Morph, levels = best.titles)
wt.morphplot$Species = "WT(cr)"

GR.morphdata = morph.data[morph.data$Genotype == "G85R(cr)",]
GR.morphdata = t(GR.morphdata[,3:8])
GR.morphdata = as.data.frame(GR.morphdata)
GR.morphdata$Morph = best.titles
GR.morphplot = data.frame(Morph = as.character(),
                          Score = as.numeric())
i=1
while(i<=9){
  temp.morph = data.frame(Morph = GR.morphdata$Morph,
                          Score = GR.morphdata[,i])
  GR.morphplot = rbind(GR.morphplot, temp.morph)
  i = i +1
}
GR.morphplot$Morph = factor(GR.morphplot$Morph, levels = best.titles)
GR.morphplot$Species = "G85R(cr)"

sum.morph = data.frame(Morph = "Total Score",
                       Score = apply(morph.data[,3:8], MARGIN = 1, sum),
                       Species = morph.data$Genotype)
sum.morph$Species = factor(sum.morph$Species, levels = c("WT(cr)", "G85R(cr)"))

all.morphplot = rbind(wt.morphplot, GR.morphplot)
all.morphplot$Species = factor(all.morphplot$Species, levels = c("WT(cr)", "G85R(cr)"))
#all.morphplot = rbind(all.morphplot, sum.morph)

ggplot(
  all.morphplot, 
  aes(x = `Score`, y = `Morph`)
  ) +
  geom_density_ridges_gradient(
    aes(fill = Species), scale = 1
    ) +
  scale_fill_manual(
    values = alpha(c("royalblue", "orange"), .5)
  ) +
  labs(title = 'Morphology Score')

ggplot(
  sum.morph, 
  aes(x = `Score`, y = `Morph`)
  ) +
  geom_density_ridges_gradient(
    aes(fill = Species), scale = 1
    ) +
  scale_fill_manual(
    values = alpha(c("royalblue", "orange"), .5)
  ) +
  labs(title = 'Morphology Score')

```{r echo = F}
all.morphplot %>% 
  ##pivot_longer(cols = -Species,
  ##             names_to = "Param",
  ##             values_to = "Value") %>% 
      ggplot(aes(x = Score, y = Morph))+
      geom_density_ridges(aes(fill = Species),
                          stat = "binline",
                          alpha = 0.5)

ggplot(
  all.morphplot, 
  aes(x = `Score`, y = `Morph`)
  ) +
  geom_density_ridges_gradient(
    #aes(fill = ..x..), scale = 3, size = 0.3
    aes(fill = Species), alpha = .9, scale = 1
    ) +
  #scale_fill_gradientn(
  #  colours = c("#0D0887FF", "#CC4678FF", "#F0F921FF"),
  #  name = "Score"
  #  )+
  labs(title = 'Morphology Score')
